# Loading and preprocessing the data
```{r}
##To set the file path.
setwd("E:\\Anoop\\Books\\Data Science\\CourseTutorials\\ReproducibleResearch\\RepData_PeerAssessment1")
###To LOad the Data set.
Data=read.csv("activity.csv",header=TRUE,sep=',',na.strings="NA")
##To remove Na vAlue from the Steps column
Data=na.omit(Dataset)
Total_steps_per_day=aggregate(steps~date,Dataset,sum)
library(ggplot2)
##Plot Histogram
qplot(Total_steps_per_day$steps,xlab='Steps',geom="histogram",ylab='Count of steps')
##barplat
##plot(x=Total_steps_per_day$day,y=Total_steps_per_day$steps,xlab='Steps',ylab='Count of steps')

```
# What is mean total number of steps taken per day?
```{r}
##Mean and Median number of steps for each Day.
Mean_Meadian=cbind(aggregate(steps~date,Data,mean)[,c(1,2)],aggregate(steps~date,Data,median)[,2])
names(Mean_Meadian)=c("Dates","Mean","Median")
Mean_Meadian
```
## The Mean and the Median on the whole are:

###Mean for the Whole Dataset is:
```{r}
Mean=mean(Data$steps)
Mean
```
###Median for the Whole Dataset is:
```{r}
Median=median(Data$steps)
Median
```


#What is the average daily activity pattern?
```{r}
Steps=aggregate(steps~interval,Data,mean)[,c(1,2)]
names(Steps)=c("Intervals","Averages")
plot(x=Steps$Intervals,y=Steps$Averages,type="l",xlab="Intervals",ylab="Averaged steps",main="Time Series Graph for Number of Steps")
```

# Imputing missing values
The Total number of missing values in the Dataset.
```{r}
Complete_data=read.csv("activity.csv",header=TRUE,sep=',',na.strings="NA")
nrow(Complete_data[complete.cases(Complete_data)==FALSE,c(1,2,3)])
```
Imputation Procedure
```{r}
 Data_without_Na=na.omit(Complete_data)
 Data_only_NA=Complete_data[complete.cases(Complete_data)==FALSE,c(1,2,3)]

##Function to calculate Mean according to Interval
imputes=function(x)
{
  Means=mean(Data_without_Na[Data_without_Na$interval==x,c(1)])
  #Means
  Data_only_NA[Data_only_NA$interval==x,c(1)]=Means
  #x
  store=rbind(Data_only_NA[Data_only_NA$interval==x,c(1,2,3)])
  }
##loop to populate across all the missing distinct intervals.
A=imputes(unique(Data_only_NA[1,c(3)]))
 
 ##loop it across all rows
for(i in  2:nrow(unique(Data_only_NA[c(3)])))
{
  row_no=i
  store=imputes(unique(Data_only_NA[i,c(3)]))
  A=rbind(A,store)
  }

# ####Completely Data with the imputed rows.

Completed_data_withimputation=rbind(Data_without_Na,A)
```

Histogram after Imputation
```{r}
Total_steps_per_day_imp=aggregate(steps~date,Completed_data_withimputation,sum)
library(ggplot2)
##Plot Histogram
qplot(Total_steps_per_day_imp$steps,xlab='Steps',geom="histogram",ylab='Count of steps')
```
##Mean and Median Daywise after Imputation
Mean and Median after Imputation
```{r}
Mean_Meadian_imp=cbind(aggregate(steps~date,Completed_data_withimputation,mean)[,c(1,2)],aggregate(steps~date,Completed_data_withimputation,median)[,2])
names(Mean_Meadian_imp)=c("Dates","Mean","Median")
Mean_Meadian_imp

```
## The Mean and the Median on the whole are:

###Mean for the Whole Dataset is:
```{r}
Mean_imp=mean(Completed_data_withimputation$steps)
Mean_imp
```
###Median for the Whole Dataset is:
```{r}
Median_imp=median(Completed_data_withimputation$steps)
Median_imp
```

##On the whole, the shape of the histogram looks the same but there are few changes,Data for few days were missing,which got added and is considered after imputation.

# Are there differences in activity patterns between weekdays and weekends?

```{r}
Completed_data_withimputation$weekdaytype=ifelse(as.POSIXlt(Completed_data_withimputation$date)$wday %in% c(1,2,3,4,5),"WeekDay","WeekEnd")
```
##Panel Plot containing Time Series Plot
```{r}
library(lattice)

Activity_data_averaged<- aggregate(steps ~ interval + weekdaytype, data=Completed_data_withimputation, mean)

xyplot(steps ~ interval | weekdaytype,Activity_data_averaged ,type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps",main='Time Series plot')
```


